---
title: "Analysis of Outer Space Objects"
subtitle: "INFO 526 - Summer 2024 - Final Project - Proposal"
author:
  - name: "Manas P Panse"
    affiliations:
      - name: "iSchool, University of Arizona"
description: "A comparison of the Outer Space Objects recorded by different entities since the 1950s."
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
  message: false
  error: false
toc: true
---

# 1 - Introduction

This project investigates and analyzes the trends and contributions of the global entities in launching outer space objects. By aiming to understand how the number of objects launched have changed over time, we can try to identify the key contributors.

# 2 - Workspace Setup

```{r}
#| label: workspace-setup
#| message: false

# Plot Theme
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14))

# Set Width of Code Output
options(width = 65)

# Set Figure Parameters for knitr
knitr::opts_chunk$set(
  fig.width = 7,        # 7" width
  fig.asp = 0.618,      # the golden ratio
  fig.retina = 3,       # dpi multiplier for displaying HTML output on retina
  fig.align = "center", # center align figures
  dpi = 300             # higher dpi, sharper image
)

# Loading Packages
pacman::p_load("dplyr", "gsubfn", "here", "knitr", "tidyr", "stringr", "tidyverse")
pacman::p_load("rworldmap", "rworldxtra", "maps", "ggplot2", "patchwork", "gtable", "gt")

# Loading Necessary Dataset
space_data <- read.csv(here("data/outer_space_objects.csv"))
gdp_data <- read.csv(here("data/gdp_data_top_5.csv"))
```

# 3 - Dataset

```{r}
#| label: dataset-info
#| message: false

glimpse(space_data)
glimpse(gdp_data)
```

I have chosen this Outer Space Objects dataset. This dataset contains detailed information about the various entities launching outer space objects over the course of time. Each entry or record contains the entity's name, the [ISO 3166-1 alpha-3 code](https://en.wikipedia.org/wiki/ISO_3166-1_alpha-3) for some specific entities, the year when this observation was captured, and the number of objects launched.

I am particularly interested in the growing field of space exploration and monitoring, understanding the trends in object launch events, and identifying these major contributors can provide valuable insights in the progress and focus areas. This analysis could potentially highlight the advancements in space technologies and the collaborative efforts of different entities worldwide.

This dataset also allows me to explore the dynamics of contributions from various countries. By analyzing these contributions over time, we can gain a deeper understanding of the shifts in focus and resource allocation in the domain of deep space monitoring. Such insights could inform strategics for future investments and collaborations in space exploration, monitoring, and soon enough, militarization.

```{r}
#| label: dataset-summary
#| message: false

space_data_summary <- paste(
  nrow(space_data), min(space_data$Year), max(space_data$Year)
)

print(space_data_summary)

gdp_data_summary <- paste(
  nrow(gdp_data)
)

print(gdp_data_summary)
```

The dataset `outer_space_objects.csv` contains 1175 entries with information on outer space objects recorded by different entities. The columns in it are : 'Entity', 'Code', 'Year', and 'num_objects'. The dataset spans from 1957, to the most recent 2023. Since this dataset is not only very comprehensive, but also primarily quantitative, it allows for a detailed analysis of trends and contributions over time.

The dataset `gdp_data_top_5.csv` contains 5 entries with information on GDPs of the BIG 5. The columns in it are : 'Country.Name', 'Country.Code', 'Indicator.Name', 'Indicator.Code', 'X2018', 'X2019', 'X2020', 'X2021', 'X2022', and 'X2023'.

# 4 - Questions

***After taking into consideration, feedback from my peers, questions have been altered.***

(1) What are the trends in the number of outer space objects launched over the years ?

(2) Which entities have contributed the most to the number of outer space objects launched ? Considering the TOP 5 Economies in the world have changed over the years, how have their space launch programs reacted to the changes in their respective economies ?

***Q3 is NOT a necessary question, but I'm just curious, so I'm going to attempt it.***

(3) If we have to predict the future, how will the the outer space object launch trend line look like ?

# 5 - Analysis Plan

***Based on my analysis, analysis plan has been altered.***

1.  Trends Over the Years :
    1.  Variables Involved - `Year`, `total_objects`.
    2.  Plan :
        -   Aggregate the space data by year.
        -   Create a smooth line plot to show the trend of total objects launched over the years.
        -   Create a line plot to show the accurate trend of total objects launched over the years, highlighting any discrepancies in trends.
        -   Highlight recent years for focus.
2.  BIG 5 :
    1.  Overall Space Launch Trends :
        -   Variables Involved - `Year`, `total_objects`, `Entity`.
        -   Plan -
            -   Aggregate the space data for the TOP 5 economies based on their GDP.
            -   Create a smooth line plot to visualize the number of objects launched by the BIG 5 since the 1950s.
            -   Use GDP data to provide context and insights into the space launch capabilities.
    2.  Geo - Distribution of Recent Launch Trends :
        -   Variables Involved - `world_map`, `Latitude`, `Longitude`, `total_objects`, `Entity`
        -   Plan -
            -   Use a world map to plot the geographic distribution of recent space launches.
            -   Overlay the launch data onto the geographic centers (centroids) of the BIG 5 respectively.
            -   Use size points to represent the total number of objects launched.
3.  Outer Space Launch Prediction :
    -   Variables Involved - `yearly_data`, `forecast_model`
    -   Plan -
        -   Implement a forecast model to predict future launches.
        -   Allow user to choose the prediction model from ARIMA, ETS, and TBATS, as well as the number of years to forecast (LIMIT 27).
        -   Display the forecast results with confidence intervals on a line plot.

# 6 - Modus Operandi

***I have something in mind for this, I just hope it's possible.***😣

***I had something in mind for this, I just hoped it's possible. IT IS AND I DID IT.*** 😎

```{r}
#| label: modus-operandi
#| echo: false

# Pulling M.O
plan <- read.csv(here("data/modusoperandi.csv"))

# Creating the gt Table with Conditional Row Styling
plan_table <- plan %>%
  gt() %>%
  tab_header(
    title = md("**Modus Operandi**"),
    subtitle = md("*Analysis of Outer Space Object Launches*")
  ) %>%
  tab_style(
    style = cell_fill(color = alpha("red", 0.5)),
    locations = cells_body(
      rows = Status == "Incomplete"
    )
  ) %>%
  tab_style(
    style = cell_fill(color = alpha("yellow", 0.5)),
    locations = cells_body(
      rows = Status == "Inprogress"
    )
  ) %>%
  tab_style(
    style = cell_fill(color = alpha("green", 0.5)),
    locations = cells_body(
      rows = Status == "Completed"
    )
  )

# Displaying M.O
plan_table
```
