---
title: "Analysis of Outer Space Objects"
subtitle: "INFO 526 - Summer 2024 - Final Project - Proposal"
author:
  - name: "Manas Panse"
    affiliations:
      - name: "iSchool, University of Arizona"
description: "A comparison of the Outer Space Objects recorded by different entities since the 1950."
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
  message: false
  error: false
toc: true
---

# 1 - Introduction

This project investigates and analyzes the trends and contributions of the global entities in recording outer space objects. By aiming to understand how the number of objects recorded have changed over time, we can identify the key contributors.

# 2 - Workspace Setup

```{r}
#| label: workspace-setup
#| message: false

# Plot theme
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14))

# set width of code output
options(width = 65)

# set figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 7,        # 7" width
  fig.asp = 0.618,      # the golden ratio
  fig.retina = 3,       # dpi multiplier for displaying HTML output on retina
  fig.align = "center", # center align figures
  dpi = 300             # higher dpi, sharper image
)

pacman::p_load("dplyr", "here", "knitr", "tidyverse")

space_data <- read.csv(here("data/outer_space_objects.csv"))
```

# 3 - Dataset

```{r}
#| label: dataset-info
#| message: false

glimpse(space_data)
```
I have chosen this Outer Space Objects dataset. This dataset contains detailed information about the various entities recording outer space objects over the course of time. Each entry or record contains the entity's name, the ISO 3166-1 alpha-3 code for some specific entities, the year when this observation was captured, and the number of objects recorded.

I am particularly interested in the growing field of space exploration and monitoring, understanding the trends in object recording, and identifying these major contributors can provide valuable insights in the progress and focus areas. This analysis could potentially highlight the advancements in space technologies and the collaborative efforts of different entities worldwide.

This dataset also allows me to explore the dynamics of contributions from various countries. By analyzing these contributions over time, we can gain a deeper understanding of the shifts in focus and resource allocation in the domain of deep space monitoring. Such insights could inform strategics for future investments and collaborations in space exploration, monitoring, and soon enough, militarization.

```{r}
#| label: dataset-summary
#| message: false

data_summary <- paste(
  nrow(space_data), min(space_data$Year), max(space_data$Year)
)

print(data_summary)
```

The dataset "outer_space_objects.csv" contains 1175 entries with information on outer space objects recorded by different entities. The columns in it are : 'Entity', 'Code', 'Year', and 'num_objects'. The dataset spans from 1957, to the most recent 2023. Since this dataset is not only very comprehensive, but also primarily quantitative, it allows for a detailed analysis of trends and contributions over time.

# 4 - Questions

(1) What are the trends in the number of outer space objects recorded over the years ?

(2) Which entities have contributed the most to the number of outer space objects recordings ? How has their contribution changed over time ?

# 5 - Analysis Plan

1. Trends Over the Years :
-   Variables Involved : '`Year`', '`num_objects`'
-   Plan :
    -   Aggregate the data by year and then sum the number of objects.
    -   Plot a time series to visualize this trend.
    
2. Contributions by Entity :
-   Variables Involved : '`Entity`', '`Year`', '`num_objects`'
-   Plan :
    -   Aggregate the data by entity and year.
    -   Plot the contributions over time to analyze changes and trends.

# 6 - Plan of Attack

+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Task Name                                  | Status     | Due     | Priority | Summary                                                                                                   |
+:==========================================:+:==========:+:=======:+:========:+:=========================================================================================================:+
| Project Proposal for Peer Review           |  complete  | 29-May  | High     | Submit the project proposal for peer review starting on May 29, 2024.                                     |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Project Proposal for Instructor Review     | incomplete | 3-Jun   | High     | Submit the project proposal for instructor review starting on June 3, 2024, after peer feedback           |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Finalize Project Outline                   | incomplete | 4-Jun   | High     | Amend the `proposal.qmd` with feedback from classmates and professor.                                     |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Aggregate Yearly Data                      | incomplete |         | High     | Aggregate the data by year to get the total number of outer space objects recorded each year.             |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Identify Top Entities                      | incomplete |         | High     | Identify the top entities contributing to the recorded outer space objects. Focus on the top 10 entities. |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Analyze Yearly Trends                      | incomplete |         | High     | Analyze the trends in the number of recorded outer space objects over the years.                          |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Entity Contribution Trends                 | incomplete |         | High     | Analyze how the contributions of the top entities have changed over time.                                 |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Include all References & Citations         | inprogress | 24-Jun  | High     | List all the references & citations used in the project.                                                  |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Finalize Report & Presentation             | incomplete | 24-Jun  | High     | Finalize the report and prepare the presentation for final submission on June 24, 2024.                   |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Visualization Techniques                   | incomplete |         | Moderate | Experiment with different visualization formats to find the best way to present the trends.               |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Draft Initial Report                       | incomplete |         | Moderate | Draft the initial report summarizing the findings and trends from the analysis.                           |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Peer Review & Feedback                     | incomplete |         | Moderate | Conduct peer review of the initial report and visualizations to gather feedback and make adjustments.     |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Prepare Conclusion and Future Work         | incomplete |         | Moderate | Summarize key insights and suggest potential areas for further research or data collection improvements.  |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Finalize the `index.qmd`                   | incomplete | 20-Jun  | Low      | Finalize the main report file `index.qmd`.                                                                |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Finalize the `presentation.qmd`            | incomplete | 20-Jun  | Low      | Finalize the presentation file `presentation.qmd`.                                                        |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+
| Record the Presentation                    | incomplete | 24-Jun  | Low      | Complete and record the presentation.                                                                     |
+--------------------------------------------+------------+---------+----------+-----------------------------------------------------------------------------------------------------------+